---
title: "NSC report"
author: "Jiahua Chen"
date: "2023-12-18"
output:
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE, fig.align = "center")
#install.packages("tidyverse")
#install.packages("modelsummary")
#install.packages("flextable")
library(officer)
library(viridis)
library(tidyverse)
library(modelsummary)
library(flextable)
```

```{r}
set_flextable_defaults(
  theme_fun = "theme_booktabs")
```

```{r}
Canada <- readxl::read_excel("code/data/Canada.xlsx")
Canada$country <- "Canada"
US <- readxl::read_excel("code/data/United States.xlsx")
US$country <- "US"
```

```{r}
df <- rbind(Canada,US)

df <- df %>%
  mutate(age = as.numeric(age)) %>% 
  mutate(birthmonth = as.numeric(birthmonth)) %>% 
  dplyr::filter(report_date > as.Date('2016-11-30')) %>% 
  dplyr::select(-c(date,details,impact,pk,direction,x,y)) %>% 
  mutate(ebike = ifelse(is.na(ebike),"NA",ebike)) %>% 
  mutate_if(is.character,as.factor) %>% 
  mutate(ebike=fct_relevel(ebike,c("Yes","No","I don't know","NA"))) 
  

```

## Data Summary

### **SInce December 2016, how many incidents involved an ebike in the U.S. and Canada?** 

NA indicates no response.

#### Raw count

```{r, echo=FALSE}
df %>% 
  mutate(country=fct_relevel(country,c("US","Canada"))) %>% 
  mutate(ebike = fct_relevel(ebike,c("Yes","No","I don't know"))) %>% 
#  dplyr::rename(Response = ebike) %>% 
  datasummary_crosstab(ebike ~ country, statistic = . ~ N, data = .,
                       output = 'flextable') %>%
  autofit() %>% 
  border_remove() %>% 
  hline_top(j = NULL, border = fp_border(color = "grey")) %>% 
  bold(bold = TRUE, part = "header") 
```

#### Percentage:

```{r}
df %>% 
  mutate(country=fct_relevel(country,c("US","Canada"))) %>% 
  mutate(ebike = fct_relevel(ebike,c("Yes","No","I don't know"))) %>% 
#  dplyr::rename(Response = ebike) %>% 
  datasummary_crosstab(ebike ~ country, statistic = . ~ Percent("col"), data = .,
                       output = 'flextable') %>% 
 autofit() %>% 
    border_remove() %>% 
  hline_top(j = NULL, border = fp_border(color = "grey")) %>% 
  bold(bold = TRUE, part = "header") 
```

#### By country

```{r}
df %>% 
 filter(country=="US") %>% 
  mutate(ebike = fct_relevel(ebike,c("Yes","No","I don't know"))) %>% 
  #dplyr::rename(Response = ebike) %>% 
  datasummary(ebike ~ N + Percent('col'), data = .,
              title = 'Ebike vs Pedal Bike in the US',
              output = 'flextable') %>% 
  autofit()   %>% 
  border_remove() %>% 
  hline_top(j = NULL, border = fp_border(color = "grey")) %>% 
  bold(bold = TRUE, part = "header") 

df %>% 
 filter(country=="Canada") %>% 
  #mutate(ebike = fct_relevel(ebike,c("Yes","No","I don't know"))) %>% 
  #dplyr::rename(ebike = ebike) %>% 
  datasummary(ebike ~ N + Percent('col'), data = .,
              title = 'Ebike vs Pedal Bike in Canada',
              output = 'flextable') %>% 
  autofit() %>% 
    border_remove() %>% 
  hline_top(j = NULL, border = fp_border(color = "grey")) %>% 
  bold(bold = TRUE, part = "header") 
```

## Gender vs Injury

```{r}
df_sum <- df %>% 
  #mutate_if(is.factor, ~replace(., . %in% values_to_convert, NA)) %>% 
     mutate(across(where(is.factor), ~fct_drop(.))) %>% #drop the changed factors
      mutate(gender = fct_collapse(gender,Male=c("['M']"),
                               Female=c("['F']"),
                               Unknown = "[]",
                               other_level = 'Other')) 
```

### **What are the gender of people who reported e-bike incidents?**

#### Raw count

```{r}
df_sum %>% 
  mutate(gender = fct_relevel(gender,c("Female","Male","Other","Unknown"))) %>% 
  datasummary(ebike ~ gender, statistic = 1~N, fmt = 0, data = .,
              output = 'flextable') %>% 
  autofit() %>% 
    border_remove() %>% 
  hline_top(j = NULL, border = fp_border(color = "grey")) %>% 
  bold(bold = TRUE, part = "header") 

```

#### Percentage

```{r}
df_sum %>% 
  mutate(gender = fct_relevel(gender,c("Female","Male","Other","Unknown"))) %>% 
  datasummary_crosstab(ebike ~ gender, statistic = . ~ Percent('row'), fmt = 1, data = .,
             # title = 'Ebike report by gender - Percentage',
              output = 'flextable') %>%
    autofit() %>% 
    border_remove() %>% 
  hline_top(j = NULL, border = fp_border(color = "grey")) %>% 
  bold(bold = TRUE, part = "header") 
```

### **What are the injury level by gender?**

```{r}
df_sum <- df_sum %>% 
  mutate(injury_level = fct_collapse(injury,
                                     "No injury"=c("No injury"),
                                     "Injury no treatment"=c("Injury, no treatment"),
                                     "Injury with treatment"=c("Injury, hospitalized",
                                                               "Injury, saw family doctor",
                                                               "Injury, hospital emergency visit"), Unknown=c("Unknown")))
```

#### Raw count

```{r}
df_sum %>% 
  mutate(injury_level = fct_relevel(injury_level,c("No injury","Injury no treatment","Injury with treatment","Unknown"))) %>% 
  filter(gender=="Female"|gender=="Male") %>% 
  droplevels() %>% 
  datasummary_crosstab(ebike*gender ~ injury_level, statistic = . ~ N, fmt = 0, data = .,
           #   title = 'Ebike injury by gender - Count',
              output = 'flextable') %>%
    autofit() %>% 
    border_remove() %>% 
  hline_top(j = NULL, border = fp_border(color = "grey")) %>% 
  bold(bold = TRUE, part = "header") 

```

#### Percentage

```{r}
df_sum %>% 
  mutate(injury_level = fct_relevel(injury_level,c("No injury","Injury no treatment","Injury with treatment","Unknown"))) %>% 
  filter(gender=="Female"|gender=="Male") %>% 
  droplevels() %>% 
  datasummary_crosstab(ebike*gender ~ injury_level, statistic = . ~ Percent('row'), fmt = 1, data = .,
             # title = 'Ebike injury by gender - Percentage',
              output = 'flextable') %>%
    autofit() %>% 
    border_remove() %>% 
  hline_top(j = NULL, border = fp_border(color = "grey")) %>% 
  bold(bold = TRUE, part = "header") 
```

#### **Graph Visualization:**

```{r}
#install.packages("viridis")
df_sum %>% 
  mutate(injury_level = fct_relevel(injury_level,c("No injury","Injury no treatment","Injury with treatment","Unknown"))) %>% 
  filter(gender=="Female"|gender=="Male") %>% 
  filter(ebike=="Yes"|ebike=="No") %>% 
  droplevels() %>% 
  ggplot(aes(x=gender,fill=injury_level)) +
  geom_bar(position="fill") +
  facet_grid(. ~ ebike) +
  scale_fill_viridis(discrete = T) +
  ylab("Percentage")
```

## Age vs Injury

```{r}
## calculate age
df_age <- df %>% mutate(month_imp = case_when(!is.na(age)&is.na(birthmonth) ~ 1,
                                              !is.na(birthmonth) ~ birthmonth)) %>%
  mutate(date = as.character(with(., sprintf("%d-%02d", age, month_imp)))) %>%
  mutate(date_imp = ifelse(is.na(age), as.integer(NA), paste(date, "-01", sep=""))) %>% 
  mutate(injury_level = fct_collapse(injury,
                                     "No injury"=c("No injury"),
                                     "Injury no treatment"=c("Injury, no treatment"),
                                     "Injury with treatment"=c("Injury, hospitalized",
                                                               "Injury, saw family doctor",
                                                               "Injury, hospital emergency visit"), Unknown=c("Unknown")))
  
df_age$date_imp <- as.Date(df_age$date_imp)

age_func <- function(dob, age.day, units = "years", floor = TRUE) {
    calc.age = lubridate::interval(dob, age.day) / lubridate::duration(num = 1, units = units)
    if (floor) return(as.integer(floor(calc.age)))
    return(calc.age)
}

df_age$age_imp <- age_func(df_age$date_imp,df_age$report_date)
df_age <- df_age %>% 
  mutate(age_group = case_when(
      age_imp <= 12            ~ "0-12",
      age_imp > 13 & age_imp <= 17 ~ "13-17",
      age_imp > 18 & age_imp <= 24 ~ "18-24",
      age_imp > 25 & age_imp <= 34 ~ "25-34",
      age_imp > 35 & age_imp <= 44 ~ "35-44",
      age_imp > 45 & age_imp <= 54 ~ "45-54",
      age_imp > 55 & age_imp <= 64 ~ "55-64",
      age_imp > 65 & age_imp <= 74 ~ "65-74",
      age_imp > 75            ~ "75+"
    ),
    age_group = as.factor(age_group)) %>% 
  mutate(age_group=fct_relevel(age_group,c("0-12","13-17","18-24","25-34","35-44","45-54","55-64","65-74","75+")))
```

### **What are the ages of people who reported e-bike incidents?**

#### Raw count

```{r}
df_age %>% 
  datasummary(ebike ~ age_group, 
                     statistic = . ~ N,data=., fmt=0,output = 'flextable') %>%
  autofit() %>% 
    border_remove() %>% 
  hline_top(j = NULL, border = fp_border(color = "grey")) %>% 
  bold(bold = TRUE, part = "header") 

```

#### Percentage

```{r}
df_age %>% 
  datasummary_crosstab(ebike ~ age_group, 
                     statistic = . ~ Percent("row"), fmt=1,data=.,
                      output = 'flextable') %>%
  autofit() %>% 
    border_remove() %>% 
  hline_top(j = NULL, border = fp_border(color = "grey")) %>% 
  bold(bold = TRUE, part = "header") 
```

#### Graph Visualization

```{r}
df_age %>% 
  filter(ebike=="Yes"|ebike=="No") %>%
  filter(!is.na(age_group)) %>% 
  droplevels() %>% 
  ggplot(aes(x=ebike,fill=age_group)) +
  geom_bar(position="fill") +
  scale_fill_viridis(discrete = T) +
  ylab("Percentage")
```

### **What are the injury level by age?**

#### **Raw count**

```{r}
df_age %>% 
  mutate(injury_level = fct_relevel(injury_level,c("No injury","Injury no treatment","Injury with treatment","Unknown"))) %>%
  filter(injury_level!="Unknown") %>% 
  filter(ebike!="I don't know"& ebike!="NA") %>% 
  droplevels() %>% 
  datasummary(ebike*injury_level ~ age_group, 
                     statistic = . ~ N,data=., fmt=0,output = 'flextable') %>%
  autofit() %>% 
    border_remove() %>% 
  hline_top(j = NULL, border = fp_border(color = "grey")) %>% 
  bold(bold = TRUE, part = "header") 

```

#### Percentage

```{r}
df_age %>% 
  filter(injury_level!="Unknown") %>% 
  filter(ebike!="I don't know"& ebike!="NA") %>% 
  droplevels() %>% 
  datasummary_crosstab(ebike*injury_level ~ age_group, 
                     statistic = . ~ Percent("row"), fmt=1,data=.,
                      output = 'flextable') %>%
  autofit() %>% 
    border_remove() %>% 
  hline_top(j = NULL, border = fp_border(color = "grey")) %>% 
  bold(bold = TRUE, part = "header") 
```

#### Graph Visualization

```{r}
df_age %>% 
  mutate(injury_level = fct_relevel(injury_level,c("No injury","Injury no treatment","Injury with treatment","Unknown"))) %>% 
  filter(!is.na(age_group)) %>% 
  filter(ebike=="Yes" | ebike=="No") %>%
  filter(injury_level!="Unknown") %>%
  droplevels() %>% 
  ggplot(aes(x=injury_level,fill=age_group)) +
  geom_bar(position="fill") +
  facet_grid(. ~ ebike) +
  scale_fill_viridis(discrete = T) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  ylab("Percentage")
```
